### 2019-04-17 2-Chapters/1-Chapter/chapter1.tex
Cette thèse me semble interminable...

@@

\begin{itemize}
\item
    \emph{GNU Radio Implementation of MALIN: ``Multi-Armed bandits Learning for Internet-of-things Networks''},
    by \textbf{Lilian Besson}, Rémi Bonnefoi \& Christophe Moy.
    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019,
    \texttt{HAL.Inria.fr/hal-02006825}.

\item
    \emph{Multi-Player Bandits Revisited},
    by \textbf{Lilian Besson} \& Emilie Kaufmann.
    \emph{Algorithmic Learning Theory} (ALT, \texttt{www.cs.cornell.edu/conferences/alt2018}), Lanzarote, Spain, April 2018,
    \texttt{HAL.Inria.fr/hal-01629733}.

\item
    \emph{Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access},\\
    by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2018.ieee-wcnc.org}), Barcelone, Spain, April 2018,
    \texttt{HAL.Inria.fr/hal-01705292}.

\item
    \emph{Multi-Armed Bandit Learning in IoT Networks and non-stationary settings},
    by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
    \emph{Conference on Cognitive Radio Oriented Wireless Networks} (CROWNCOM, \texttt{crowncom.org/2017}), Lisboa, Portugal, Septembre $2017$,
    \texttt{HAL.Inria.fr/hal-01575419},
    \textbf{Best Paper Award}.

@@

\item
    \emph{Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions},\\
    by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
    Presented at the 1st International MOTIoN workshop (\texttt{sites.google.com/view/wcncworkshop-motion2019}),
    at \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019,

\item
    \emph{MALIN: ``Multi-Arm bandit Learning for Iot Networks'' with GRC: A TestBed Implementation and Demonstration that Learning Helps},
    by \textbf{Lilian Besson}, Rémi Bonnefoi, Christophe Moy.
    Demonstration presented at \emph{International Conference on Communication} (ICT, \texttt{ict-2018.org/demos}), in Saint-Malo, France in June $2018$.
    See \texttt{YouTu.be/HospLNQhcMk} for a $6$-minutes presentation video.

@@

\begin{itemize}
\item
    \emph{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires} (in French),
    submitted at GRETSI 2019, \texttt{GRETSI.fr/colloque2019},
    by \textbf{Lilian Besson} \& Emilie Kaufmann, March $2019$,
    \texttt{HAL.Inria.fr/hal-02006471}.

\item
    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
    submitted at COLT 2019, \texttt{LearningTheory.org/colt2019/},
    by \textbf{Lilian Besson} \& Emilie Kaufmann, February $2019$,
    \texttt{HAL.Inria.fr/hal-02006471}.

@@

\begin{itemize}
\item
    \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
    by \textbf{Lilian Besson}, active development since October $2016$,
    \texttt{HAL.Inria.fr/hal-01840022}.
    About $40000$ lines of code, hosted on \texttt{GitHub.com/SMPyBandits},
    with a complete documentation on \texttt{SMPyBandits.rtfd.io}.

\item
    \emph{What Doubling-Trick Can and Can't Do for Multi-Armed Bandits},\\
    by \textbf{Lilian Besson} \& Emilie Kaufmann, September $2018$,
    \texttt{HAL.Inria.fr/hal-01736357}.

>>>>

\TODOL{Write this!}

@@

\begin{itemize}
\item
    \emph{GNU Radio Implementation of MALIN: ``Multi-Armed bandits Learning for Internet-of-things Networks''},
    by \textbf{Lilian Besson}, Rémi Bonnefoi \& Christophe Moy.
    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019,
    \texttt{HAL.Inria.fr/hal-02006825}.
    \cite{Besson2019WCNC}

\item
    \emph{Multi-Player Bandits Revisited},
    by \textbf{Lilian Besson} \& Emilie Kaufmann.
    \emph{Algorithmic Learning Theory} (ALT, \texttt{www.cs.cornell.edu/conferences/alt2018}), Lanzarote, Spain, April 2018,
    \texttt{HAL.Inria.fr/hal-01629733}.
    \cite{Besson2018ALT}

\item
    \emph{Aggregation of Multi-Armed Bandits learning algorithms for Opportunistic Spectrum Access},\\
    by \textbf{Lilian Besson}, Emilie Kaufmann \& Christophe Moy.
    \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2018.ieee-wcnc.org}), Barcelone, Spain, April 2018,
    \texttt{HAL.Inria.fr/hal-01705292}.
    \cite{Besson2018WCNC}

\item
    \emph{Multi-Armed Bandit Learning in IoT Networks and non-stationary settings},
    by Rémi Bonnefoi, \textbf{Lilian Besson}, Christophe Moy, Emilie Kaufmann \& J. Palicot.
    \emph{Conference on Cognitive Radio Oriented Wireless Networks} (CROWNCOM, \texttt{crowncom.org/2017}), Lisboa, Portugal, Septembre $2017$,
    \texttt{HAL.Inria.fr/hal-01575419},
    \textbf{Best Paper Award}.
    \cite{Bonnefoi17}

@@

\item
    \emph{Upper-Confidence Bound for Channel Selection in LPWA Networks with Retransmissions},\\
    by Rémi Bonnefoi, \textbf{Lilian Besson}, J. Manco-Vasquez \& Christophe Moy.
    1st International MOTIoN workshop (\texttt{sites.google.com/view/wcncworkshop-motion2019}),
    at \emph{Wireless Communication and Networks Conference} (IEEE WCNC, \texttt{wcnc2019.ieee-wcnc.org}), Marrakech, Morocco, April 2019.
    \cite{Bonnefoi2019WCNC}

\item
    \emph{MALIN: ``Multi-Arm bandit Learning for Iot Networks'' with GRC: A TestBed Implementation and Demonstration that Learning Helps},
    by \textbf{Lilian Besson}, Rémi Bonnefoi, Christophe Moy.
    Demonstration presented at \emph{International Conference on Communication} (ICT, \texttt{ict-2018.org/demos}), in Saint-Malo, France in June $2018$.
    See \texttt{YouTu.be/HospLNQhcMk} for a $6$-minutes presentation video.
    \cite{Besson2018ICT}

@@

\begin{itemize}
\item
    \emph{Analyse non asymptotique d'un test séquentiel de détection de ruptures et application aux bandits non stationnaires} (in French),
    submitted at GRETSI 2019, \texttt{GRETSI.fr/colloque2019},
    by \textbf{Lilian Besson} \& Emilie Kaufmann, March $2019$,
    \texttt{HAL.Inria.fr/hal-02006471}.
    \cite{Besson2019Gretsi}

\item
    \emph{The Generalized Likelihood Ratio Test meets klUCB: an Improved Algorithm for Piece-Wise Non-Stationary Bandits},
    submitted at COLT 2019, \texttt{LearningTheory.org/colt2019/},
    by \textbf{Lilian Besson} \& Emilie Kaufmann, February $2019$,
    \texttt{HAL.Inria.fr/hal-02006471}.
    \cite{Besson2019GLRT}

@@

\begin{itemize}
\item
    \emph{SMPyBandits: an Open-Source Research Framework for Single and Multi-Players Multi-Arms Bandits (MAB) Algorithms in Python},
    by \textbf{Lilian Besson}, active development since October $2016$,
    \texttt{HAL.Inria.fr/hal-01840022}.
    About $40000$ lines of code, hosted on \texttt{GitHub.com/SMPyBandits},
    with a complete documentation on \texttt{SMPyBandits.rtfd.io}.
    \cite{SMPyBandits,SMPyBanditsJMLR}

\item
    \emph{What Doubling-Trick Can and Can't Do for Multi-Armed Bandits},\\
    by \textbf{Lilian Besson} \& Emilie Kaufmann, September $2018$,
    \texttt{HAL.Inria.fr/hal-01736357}.
    \cite{Besson2018DoublingTricks}

### 2019-04-17 2-Chapters/3-Chapter/chapter3.tex
Expérience en cours pour la fin du Chapitre:3.
Done for today...
Done for this simulation
Cette thèse me semble interminable...

% https://tex.stackexchange.com/a/12430/
\begin{small}
    \begin{listing}[h!]
        \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
# 1. get the code in the folder you want
$ git clone https://GitHub.com/SMPyBandits/SMPyBandits.git
$ cd SMPyBandits.git
# 2. install the requirements
$ pip install -r requirements.txt
        \end{minted}
        \caption{Small snippet of Bash to download and install dependencies of SMPyBandits}
        % \captionof{listing}{Small snippet of Bash to download and install dependencies of SMPyBandits \label{lst:3:howToInstallLibrary}.}
        \label{lst:3:howToInstallLibrary}
    \end{listing}
\end{small}

@@

% https://tex.stackexchange.com/a/12430/
\begin{small}
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
# 3. run a single-player simulation
$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=4 \
  MEANS=[0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9] \
  python3 main.py configuration.py
    \end{minted}
    \caption{Small snippet of Bash to run a simple experiment with SMPyBandits}
    % \captionof{listing}{Small snippet of Bash to run a simple experiment with SMPyBandits \label{lst:3:howToRunBasicLibrary}.}
    \label{lst:3:howToRunBasicLibrary}
\end{listing}
\end{small}

@@

% https://tex.stackexchange.com/a/12430/
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
$ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
    \end{minted}
    \caption{Small snippet of Bash to run an experiment}
    % \captionof{listing}{Small snippet of Bash to run an experiment \label{lst:3:howToRunExperiment2}.}
    \label{lst:3:howToRunExperiment2}
\end{listing}

@@

\TODOL{Je veux terminer cet exemple!}

% FIXME run
% for N in 1 10 100; do for NJOBS in 1 4 100; do DEBUG=True SAVEALL=False NOPLOTS=True N=$N T=10000 N_JOBS=$NJOBS make single; read; done

\begin{table}[ht]
% \begin{footnotesize}
    \centering
    \begin{tabular}{c|ccc}
    \textbf{Repetitions} $\;$ \textbackslash $\;$ \texttt{N\_JOBS} & $1$ & $4$ & $20$ \\
        \hline
        $1$ repetition    & $XXX \;\text{s}$ & $XXX \;\text{s}$ & $XXX \;\text{s}$ \\
        $10$ repetitions  & $XXX \;\text{s}$ & $XXX \;\text{s}$ & $XXX \;\text{s}$ \\
        $100$ repetitions & $59 \;\text{s}$ & $37 \;\text{s}$ & $XXX \;\text{s}$ \\
        \hline
    \end{tabular}
    \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$.}
    \label{table:3:speedUpTimeParallelComputations}
% \end{footnotesize}
\end{table}

>>>>

% https://tex.stackexchange.com/a/12430/
\begin{small}
    \begin{listing}[h!]
        \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
# 1. get the code in the folder you want
$ git clone https://GitHub.com/SMPyBandits/SMPyBandits.git
$ cd SMPyBandits.git
# 2. install the requirements
$ pip install -r requirements.txt
        \end{minted}
        \caption{Small snippet of Bash code to download and install dependencies of SMPyBandits}
        % \captionof{listing}{Small snippet of Bash code to download and install dependencies of SMPyBandits \label{lst:3:howToInstallLibrary}.}
        \label{lst:3:howToInstallLibrary}
    \end{listing}
\end{small}

@@

% https://tex.stackexchange.com/a/12430/
\begin{small}
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
# 3. run a single-player simulation
$ BAYES=False ARM_TYPE=Bernoulli N=1000 T=10000 K=9 N_JOBS=4 \
  MEANS=[0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9] \
  python3 main.py configuration.py
    \end{minted}
    \caption{Small snippet of Bash code to run a simple experiment with SMPyBandits}
    % \captionof{listing}{Small snippet of Bash code to run a simple experiment with SMPyBandits \label{lst:3:howToRunBasicLibrary}.}
    \label{lst:3:howToRunBasicLibrary}
\end{listing}
\end{small}

@@

% https://tex.stackexchange.com/a/12430/
\begin{listing}[h!]
    \begin{minted}[linenos=true,numbersep=5pt,frame=lines,framesep=2mm]{bash}
$ T=50000 N=100 K=10 ARM_TYPE=Gaussian python main.py
    \end{minted}
    \caption{Small snippet of Bash code to run an experiment}
    % \captionof{listing}{Small snippet of Bash code to run an experiment \label{lst:3:howToRunExperiment2}.}
    \label{lst:3:howToRunExperiment2}
\end{listing}

@@

% for N in 1 10 100; do for NJOBS in 1 4 100; do DEBUG=True SAVEALL=False NOPLOTS=True N=$N T=10000 N_JOBS=$NJOBS make single; read; done

\begin{table}[ht]
% \begin{footnotesize}
    \centering
    \begin{tabular}{c|ccc}
    \textbf{Repetitions} $\;$ \textbackslash $\;$ \texttt{N\_JOBS} & $1$ & $4$ & $20$ \\
        \hline
        $1$ repetition    & $15 \;\text{s}$ & $26 \;\text{s}$ & $43 \;\text{s}$ \\
        $10$ repetitions  & $87 \;\text{s}$ & $51 \;\text{s}$ & $76 \;\text{s}$ \\
        $100$ repetitions & $749 \;\text{s}$ & $272 \;\text{s}$ & $308 \;\text{s}$ \\
        $500$ repetitions & $2944 \;\text{s}$ & $1530 \;\text{s}$ & $1346 \;\text{s}$ \\
        \hline
    \end{tabular}
    \caption{Effect on the running time of using \texttt{N\_JOBS} jobs in parallel, for a simulation with $9$ different algorithms, for $K=9$ arms, a time horizon of $T=10000$.}
    \label{table:3:speedUpTimeParallelComputations}
% \end{footnotesize}
\end{table}

### 2019-04-17 3-Appendices/1-Appendix/appendix1.tex
Ok for tex root
WOrking on the whole document but not yet Chapter5
Cette thèse me semble interminable...

%!TEX root = ../thesis.tex
% ******************************* Thesis Appendix A ****************************
\chapter{Appendix for Chapter \ref{chapter:1}}
\label{app:chapt1}

In this appendix, blabla.

>>>>

%!TEX root = ../PhD_thesis__Lilian_Besson
% ******************************* Thesis Appendix A ****************************
\chapter{Additional Mathematical Tools}
\label{app:1:mathsTools}

In this appendix, we present some mathematical tools that are used in various proofs in this thesis.

% ----------------------------------------------------------------------------
\section{Concentration Inequalities}
\label{sec:app1:ConcentrationInequalities}

\TODOL{I probably need to include Hoeffding's inequality, Chernoff's method etc?}

Reference: \cite{Boucheron2013} and maybe the original paper by Hoeffding, Chernoff etc.

% ----------------------------------------------------------------------------
\section{Useful lemmas}
\label{sec:app1:UsefulLemmas}

\TODOL{I probably need to include a bunch of useful lemmas, used in various places in the thesis, like Pinsker's inequality, etc.}

% ----------------------------------------------------------------------------
\section{Results about klUCB}
\label{sec:app1:resultsklUCB}

\TODOL{I probably need to include some results about klUCB, mainly coming from \cite{KLUCBJournal}, used in various places in the thesis?}

### 2019-04-17 3-Appendices/2-Appendix/appendix2.tex
Cette thèse me semble interminable...

%!TEX root = ../PhD_thesis__Lilian_Besson
% ******************************* Thesis Appendix B ****************************
\chapter{About another publication: ``What Doubling-Trick Can and Can't Do for Multi-Armed Bandits''}
\label{app:2:DoublingTricks}

This appendix quickly presents the contributions of another publication that was not presented in the main text of this thesis, \cite{Besson2018DoublingTricks}.

% \TODOL{
% I cannot include it raw without more work on the maths, we have to write only the best versions of our results.

% But I am interested in including this article in my thesis, mainly because GLRklUCB in Chapter 6 is not anytime!

% I need to try DoublingTrick + GLRklUCB, at least experimentally!
% For the theoretical analysis, it will probably be a huge failure.
% }

@@

A well-known technique to obtain an anytime algorithm from any non-anytime algorithm is the ``Doubling Trick'', as first introduced in \cite{CesaLugosi06}.
We studied doubling tricks between fall 2017 and spring 2018, and we wrote an article \cite{Besson2018DoublingTricks} which was rejected at the COLT 2018 conference, and unfortunately we lacked the time to improve it and resubmit it to another conference.
%
In the context of adversarial or stochastic multi-armed bandits,
the performance of an algorithm is measured by its regret,
and we study two families of sequences of growing horizons (geometric and exponential)
to generalize previously known results that certain doubling tricks can be used to conserve certain regret bounds.
In a broad setting, we prove that a geometric doubling trick can be used to conserve (minimax) bounds in $R_T = \cO(\sqrt{T})$ but \emph{cannot} conserve (distribution-dependent) bounds in $R_T = \cO(\log T)$.
We give insights as to why exponential doubling tricks may be better, as they conserve bounds in $R_T = \cO(\log T)$, and are close to conserving bounds in $R_T = \cO(\sqrt{T})$.
Interestingly, we prove that they conserve bounds of the mixed form $R_T = \cO(T^{\gamma} (\log(T))^{\delta})$, for $0<\gamma<1$ and $\delta>0$, and so an exponential doubling trick could be used to obtain an anytime version of the algorithm we propose for non-stationary bandits, \GLRklUCB{} in Chapter~\ref{chapter:6}.
However, our study was only focussing on stationary bandit, and it is left as a future work to explore the harder case of piece-wise stationary problems.

>>>>

%!TEX root = ../PhD_thesis__Lilian_Besson
% ******************************* Thesis Appendix B ****************************
% \chapter{About another publication: ``What Doubling-Trick Can and Can't Do for Multi-Armed Bandits''}
\chapter{About Doubling-Trick for Multi-Armed Bandits}
\label{app:2:DoublingTricks}

This appendix quickly presents the contributions of another publication that was not presented in the main text of this thesis.
We studied doubling tricks and their possible use for multi-armed bandits, between fall 2017 and spring 2018, and we wrote an article \cite{Besson2018DoublingTricks} which was rejected at the COLT 2018 conference, and unfortunately we lacked the time to improve it and resubmit it to another conference.

\TODOL{
I cannot include it raw without more work on the maths, we have to write only the best versions of our results.

But I am interested in including this article in my thesis, mainly because GLRklUCB in Chapter 6 is not anytime!

I need to try DoublingTrick + GLRklUCB, at least experimentally!
For the theoretical analysis, it will probably be a huge failure.
}

@@

A well-known technique to obtain an anytime algorithm from any non-anytime algorithm is the ``Doubling Trick'', as first introduced in \cite{CesaLugosi06}.
%
In the context of adversarial or stochastic multi-armed bandits,
the performance of an algorithm is measured by its regret,
and we study two families of sequences of growing horizons (geometric and exponential)
to generalize previously known results that certain doubling tricks can be used to conserve certain regret bounds.
In a broad setting, we prove that a geometric doubling trick can be used to conserve (minimax) bounds in $R_T = \cO(\sqrt{T})$ but \emph{cannot} conserve (distribution-dependent) bounds in $R_T = \cO(\log T)$.
We give insights as to why exponential doubling tricks may be better, as they conserve bounds in $R_T = \cO(\log T)$, and are close to conserving bounds in $R_T = \cO(\sqrt{T})$.
Interestingly, we prove that they conserve bounds of the mixed form $R_T = \cO(T^{\gamma} (\log(T))^{\delta})$, for $0<\gamma<1$ and $\delta>0$, and so an exponential doubling trick could be used to obtain an anytime version of the algorithm we propose for non-stationary bandits, \GLRklUCB{} in Chapter~\ref{chapter:6}.
However, our study was only focussing on stationary bandit, and it is left as a future work to explore the harder case of piece-wise stationary problems.
